Cultural evolutionary dynamics with prestige bias

### 2022-12-17 ms/ms.tex
update

\subsection*{Dichotomous trait}
To allow for mathematical analysis of the model, we introduce a simplified version where the trait only has two phenotypes: the optimal phenotype $\hat{A}$ and the sub-optimal phenotype $A$. 
All role-models with the same phenotype will contribute to the probability of that phenotype to be transmitted, and thus influence is determined by the number of copiers that have already chosen a role-model with either phenotype.
In addition, we assume homogeneous $\alpha$ for simplicity and for easier mathematical analysis.
Therefore, the probability of the $i$-th copier to copy phenotype $A$ is
\begin{equation}\label{eq:binary-model}
G_{i,A} = \frac{(N-C)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-C)\alpha'\beta(A) + C\alpha'\beta(\hat{A})} = \frac{(N-C)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-C)\alpha'\beta(A) + \alpha'C}
\end{equation}
where $C$ is the number of role-models with trait $\hat{A}$ and $K_{i,A}$ is the number of copiers that already chose $A$ when copier $i$ chooses a role-model and $\alpha'$ is the odd ratio of the bias weight $\alpha'= \frac{\alpha}{1-\alpha}$.

@@

>>>>

\subsection*{Dichotomous trait}
To allow for mathematical analysis of the model, we introduce a simplified version where the trait only has two phenotypes: the optimal phenotype $\hat{A}$ and the sub-optimal phenotype $A$. 
All role-models with the same phenotype will contribute to the probability of that phenotype to be transmitted, and thus influence is determined by the number of copiers that have already chosen a role-model with either phenotype.
In addition, we assume homogeneous $\alpha$, also called \textit{exchangability}, for simplicity and for easier mathematical analysis.
Therefore, the probability of the $i$-th copier to copy phenotype $A$ is
\begin{equation}\label{eq:binary-model}
G_{i,A} = \frac{(N-C)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-C)\alpha'\beta(A) + C\alpha'\beta(\hat{A})} = \frac{(N-C)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-C)\alpha'\beta(A) + \alpha'C}
\end{equation}
where $C$ is the number of role-models with trait $\hat{A}$ and $K_{i,A}$ is the number of copiers that already chose $A$ when copier $i$ chooses a role-model and $\alpha'$ is the odd ratio of the bias weight $\alpha'= \frac{\alpha}{1-\alpha}$.

@@

So far we examined different variations of our model using a Wright-Fisher model as our basis, mainly because we wanted to base our model on a known one from the literature, namely \citet{evolutionBook} indirect bias model. When approximating our model by comparing it to a Polya Urn process, we noticed that our model would be identical to a model where every copier that chose a role-model would copy its trait and join the role-models during the selection process. Based on this, we theorize that when using a Moran model as basis instead of a Wright-Fisher mode, we would get very similar, and perhaps identical results. It would be interesting to run simulations and analyze that model, to prove or disprove our hypothesis.
